---
title: "Computational Statistics in Business Analytics - Confidence Intervals"
author: "Daniele Melotti"
date: "Nov 2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data Standardization and Exploration

## Standardize a synthetic dataset

Data standardization is essentially the process of rescaling data so that it has a mean of 0 and a standard deviation of 1. This process is particularly helpful in guaranteeing data consistency, allowing for comparisons between variables that are measured with different scales.
To execute data standardization, we need to subtract the mean of a vector from all its values, and divide this difference by the standard deviation to get a vector of standardized values. We demonstrate this on a synthetic dataset, which we create following a normal distribution with mean of 940 and standard deviation of 190.
```{r}
# Setting seed
set.seed(194)

rnorm_raw <- rnorm(n = 100, mean = 940, sd = 190) # non-standardized data
rnorm_std <- (rnorm_raw - mean(rnorm_raw))/sd(rnorm_raw) # standardized data
```

## Analyze the standardized data

We shall make sure that the mean and standard deviation of the standardized dataset are 0 and 1 respectively:
```{r}
mean(rnorm_std)
sd(rnorm_std)
```

This is the case indeed. **What shape of the standardized distribution do we expect in comparison to that of the non standardized data?** The shape of the distribution of rnorm_std should be bell-shaped, as for a normal distribution. That’s because _the standardization process does not transform the underlying distribution structure of the data_ (hence, it does not influence the shape of the distribution); standardization only turns raw data into z-scores (it changes the scale of data), therefore, if the non-standardized distribution is normal (bell-shaped), its standardized version will also be normal. Similarly, a non-normally shaped distribution will remain non-normally shaped after standardization. Let’s confirm this:
```{r}
par(mfrow = c(1, 2))
plot(density(rnorm_raw), lwd = 2, main = "Non Standardized Distribution", col = "tomato")
plot(density(rnorm_std), lwd = 2, main = "Standardized Distribution", col = "cornflowerblue")
```

Indeed, the shape of both distributions is identical and roughly bell-shaped.

# 2. Confidence Intervals Simulation


